

SBS5242
Synthetic subset of the Austrian structural business statistics data

#### Description 

Synthetic subset of the Austrian structural business statistics (SBS) data, namely NACE code 52.42
(retail sale of clothing).scattJitt
55

#### Details 

The Austrian SBS data set consists of more than 320.000 enterprises. Available raw (unedited) data
set: 21669 observations in 90 variables, structured according NACE revision 1.1 with 3891 missing
values.
We investigate 9 variables of NACE 52.42 (retail sale of clothing).
From these confidential raw data set a non-confidential, close-to-reality, synthetic data set was gen-
erated.
Source
http://www.statistik.at

#### Examples 
```{r}
data(SBS5242)
aggr(SBS5242)
scattJitt
Bivariate jitter plot

#### Description 

Create a bivariate jitter plot.

#### Usage 
<pre><code>

scattJitt(x, delimiter = NULL, col = c("skyblue", "red", "red4", "orange",
"orange4"), alpha = NULL, cex = par("cex"), col.line = "lightgrey",
lty = "dashed", lwd = par("lwd"), numbers = TRUE,
cex.numbers = par("cex"), main = NULL, sub = NULL, xlab = NULL,
ylab = NULL, axes = TRUE, frame.plot = axes, labels = c("observed",
"missing", "imputed"), ...)
Arguments
x a data.frame or matrix with two columns.
delimiter a character-vector to distinguish between variables and imputation-indices for
imputed variables (therefore, x needs to have colnames). If given, it is used
to determine the corresponding imputation-index for any imputed variable (a
logical-vector indicating which values of the variable have been imputed). If
such imputation-indices are found, they are used for highlighting and the colors
are adjusted according to the given colors for imputed variables (see col).56
scattJitt
col
alpha
cex
col.line
lty
lwd
numbers
cex.numbers
main, sub
xlab, ylab
axes
frame.plot
labels
...
a vector of length five giving the colors to be used in the plot. The first color will
be used for complete observations, the second/fourth color for missing/imputed
values in only one variable, and the third/fifth color for missing/imputed values
in both variables. If only one color is supplied, it is used for all. Else if two
colors are supplied, the second one is recycled.
a numeric value between 0 and 1 giving the level of transparency of the colors,
or NULL. This can be used to prevent overplotting.
the character expansion factor for the plot characters.
the color for the lines dividing the plot region.
the line type for the lines dividing the plot region (see par).
the line width for the lines dividing the plot region.
a logical indicating whether the frequencies of observed and missing/imputed
values should be displayed (see ‘
#### Details 
’).
the character expansion factor to be used for the frequencies of the observed and
missing/imputed values.
main and sub title.
axis labels.
a logical indicating whether both axes should be drawn on the plot. Use graphi-
cal parameter "xaxt" or "yaxt" to suppress just one of the axes.
a logical indicating whether a box should be drawn around the plot.
a vector of length three giving the axis labels for the regions for observed, miss-
ing and imputed values (see ‘
#### Details 
’).
further graphical parameters to be passed down (see par).

#### Details 

The amount of observed and missing/imputed values is visualized by jittered points. Thereby the plot region is divided into up to four regions according to the existence of missing/imputed values
in one or both variables. In addition, the amount of observed and missing/imputed values can be represented by a number.
Note
Some of the argument names and positions have changed with version 1.3 due to extended function-
ality and for more consistency with other plot functions in VIM. For back compatibility, the argument cex.text can still be supplied to ...{} and is handled correctly. Nevertheless, it is deprecated and
no longer documented. Use cex.numbers instead.
Author(s)
Matthias Templ, modifications by Andreas Alfons and Bernd Prantner
References
M. Templ, A. Alfons, P. Filzmoser (2012) Exploring incomplete data using visualization tools.
Journal of Advances in Data Analysis and Classification, Online first. DOI: 10.1007/s11634-011-
0102-y.scattmatrixMiss
57

#### Examples 
```{r}
data(tao, package = "VIM")
## for missing values
scattJitt(tao[, c("Air.Temp", "Humidity")])
## for imputed values
scattJitt(kNN(tao[, c("Air.Temp", "Humidity")]), delimiter = "_imp")
scattmatrixMiss
Scatterplot matrix with information about missing/imputed values

#### Description 

Scatterplot matrix in which observations with missing/imputed values in certain variables are high-
lighted.

#### Usage 
<pre><code>

scattmatrixMiss(x, delimiter = NULL, highlight = NULL,
selection = c("any", "all"), plotvars = NULL, col = c("skyblue", "red",
"orange"), alpha = NULL, pch = c(1, 3), lty = par("lty"),
diagonal = c("density", "none"), interactive = TRUE, ...)
Arguments
x a matrix or data.frame.
delimiter a character-vector to distinguish between variables and imputation-indices for imputed variables (therefore, x needs to have colnames). If given, it is used
to determine the corresponding imputation-index for any imputed variable (a
logical-vector indicating which values of the variable have been imputed). If such imputation-indices are found, they are used for highlighting and the colors
are adjusted according to the given colors for imputed variables (see col).
highlight a vector giving the variables to be used for highlighting. If NULL (the default),
all variables are used for highlighting.
selection the selection method for highlighting missing/imputed values in multiple high-
light variables. Possible values are "any" (highlighting of missing/imputed values in any of the highlight variables) and "all" (highlighting of missing/imputed
values in all of the highlight variables).
plotvars a vector giving the variables to be plotted. If NULL (the default), all variables are plotted.
col a vector of length three giving the colors to be used in the plot. The second/third color will be used for highlighting missing/imputed values.
alpha a numeric value between 0 and 1 giving the level of transparency of the colors, or NULL. This can be used to prevent overplotting.58
scattmatrixMiss
pch a vector of length two giving the plot characters. The second plot character will
be used for the highlighted observations.
lty a vector of length two giving the line types for the density plots in the diagonal panels (if diagonal="density"). The second line type is used for the
highlighted observations. If a single value is supplied, it is used for both non-
highlighted and highlighted observations.
diagonal a character string specifying the plot to be drawn in the diagonal panels. Pos-
sible values are "density" (density plots for non-highlighted and highlighted
observations) and "none".
interactive a logical indicating whether the variables to be used for highlighting can be selected interactively (see ‘
#### Details 
’).
... for scattmatrixMiss, further arguments and graphical parameters to be passed
to pairsVIM. par("oma") will be set appropriately unless supplied (see par).
For TKRscattmatrixMiss, further arguments to be passed to scattmatrixMiss.

#### Details 

scattmatrixMiss uses pairsVIM with a panel function that allows highlighting of missing/imputed
values.
If interactive=TRUE, the variables to be used for highlighting can be selected interactively. Ob-
servations with missing/imputed values in any or in all of the selected variables are highlighted (as
determined by selection). A variable can be added to the selection by clicking in a diagonal panel.
If a variable is already selected, clicking on the corresponding diagonal panel removes it from the
selection. Clicking anywhere else quits the interactive session.
The graphical parameter oma will be set unless supplied as an argument.
TKRscattmatrixMiss behaves like scattmatrixMiss, but uses tkrplot to embed the plot in a
Tcl/Tk window. This is useful if the number of variables is large, because scrollbars allow to move
from one part of the plot to another.
Note
Some of the argument names and positions have changed with version 1.3 due to a re-implementation
and for more consistency with other plot functions in VIM. For back compatibility, the argument
colcomb can still be supplied to ...{} and is handled correctly. Nevertheless, it is deprecated and
no longer documented. Use highlight instead. The arguments smooth, reg.line and legend.plot
are no longer used and ignored if supplied.
Author(s)
Andreas Alfons, Matthias Templ, modifications by Bernd Prantner
References
M. Templ, A. Alfons, P. Filzmoser (2012) Exploring incomplete data using visualization tools.
Journal of Advances in Data Analysis and Classification, Online first. DOI: 10.1007/s11634-011-
0102-y.scattMiss
59
See Also
pairsVIM, marginmatrix

#### Examples 
```{r}
data(sleep, package = "VIM")
## for missing values
x <- sleep[, 1:5]
x[,c(1,2,4)] <- log10(x[,c(1,2,4)])
scattmatrixMiss(x, highlight = "Dream")
## for imputed values
x_imp <- kNN(sleep[, 1:5])
x_imp[,c(1,2,4)] <- log10(x_imp[,c(1,2,4)])
scattmatrixMiss(x_imp, delimiter = "_imp", highlight = "Dream")
scattMiss
Scatterplot with information about missing/imputed values

#### Description 

In addition to a standard scatterplot, lines are plotted for the missing values in one variable. If there
are imputed values, they will be highlighted.

#### Usage 
<pre><code>

scattMiss(x, delimiter = NULL, side = 1, col = c("skyblue", "red",
"orange", "lightgrey"), alpha = NULL, lty = c("dashed", "dotted"),
lwd = par("lwd"), quantiles = c(0.5, 0.975), inEllipse = FALSE,
zeros = FALSE, xlim = NULL, ylim = NULL, main = NULL, sub = NULL,
xlab = NULL, ylab = NULL, interactive = TRUE, ...)
Arguments
x a matrix or data.frame with two columns.
delimiter a character-vector to distinguish between variables and imputation-indices for
imputed variables (therefore, x needs to have colnames). If given, it is used
to determine the corresponding imputation-index for any imputed variable (a
logical-vector indicating which values of the variable have been imputed). If
such imputation-indices are found, they are used for highlighting and the colors
are adjusted according to the given colors for imputed variables (see col).
side if side=1, a rug representation and vertical lines are plotted for the missing/imputed
values in the second variable; if side=2, a rug representation and horizontal lines
for the missing/imputed values in the first variable.60
scattMiss
col a vector of length four giving the colors to be used in the plot. The first color
is used for the scatterplot, the second/third color for the rug representation for
missing/imputed values. The second color is also used for the lines for missing
values. Imputed values will be highlighted with the third color, and the fourth
color is used for the ellipses (see ‘
#### Details 
’). If only one color is supplied, it is
used for the scatterplot, the rug representation and the lines, whereas the default
color is used for the ellipses. Else if a vector of length two is supplied, the
default color is used for the ellipses as well.
alpha a numeric value between 0 and 1 giving the level of transparency of the colors,
or NULL. This can be used to prevent overplotting.
lty a vector of length two giving the line types for the lines and ellipses. If a single
value is supplied, it will be used for both.
lwd a vector of length two giving the line widths for the lines and ellipses. If a single
value is supplied, it will be used for both.
quantiles a vector giving the quantiles of the chi-square distribution to be used for the
tolerance ellipses, or NULL to suppress plotting ellipses (see ‘
#### Details 
’).
inEllipse plot lines only inside the largest ellipse. Ignored if quantiles is NULL or if there
are imputed values.
zeros a logical vector of length two indicating whether the variables are semi-continuous,
i.e., contain a considerable amount of zeros. If TRUE, only the non-zero observa-
tions are used for computing the tolerance ellipses. If a single logical is supplied,
it is recycled. Ignored if quantiles is NULL.
xlim, ylim axis limits.
main, sub main and sub title.
xlab, ylab axis labels.
interactive a logical indicating whether the side argument can be changed interactively ((see details)..
... further graphical parameters to be passed down (see par).

#### Details 

Information about missing values in one variable is included as vertical or horizontal lines, as deter-
mined by the side argument. The lines are thereby drawn at the observed x- or y-value. In case of
imputed values, they will additionally be highlighted in the scatterplot. Supplementary, percentage
coverage ellipses can be drawn to give a clue about the shape of the bivariate data distribution.
If interactiveis TRUE, clicking in the bottom margin redraws the plot with information about
missing/imputed values in the first variable and clicking in the left margin redraws the plot with
information about missing/imputed values in the second variable. Clicking anywhere else in the
plot quits the interactive session.
Note
The argument zeros has been introduced in version 1.4. As a result, some of the argument positions
have changed.sleep
61
Author(s)
Andreas Alfons, modifications by Bernd Prantner
References
M. Templ, A. Alfons, P. Filzmoser (2012) Exploring incomplete data using visualization tools.
Journal of Advances in Data Analysis and Classification, Online first. DOI: 10.1007/s11634-011-
0102-y.
See Also
marginplot

#### Examples 
```{r}
data(tao, package = "VIM")
## for missing values
scattMiss(tao[,c("Air.Temp", "Humidity")])
## for imputed values
scattMiss(kNN(tao[,c("Air.Temp", "Humidity")]), delimiter = "_imp")
sleep
Mammal sleep data

#### Description 

Sleep data with missing values.
Format
A data frame with 62 observations on the following 10 variables.
BodyWgt a numeric vector
BrainWgt a numeric vector
NonD a numeric vector
Dream a numeric vector
Sleep a numeric vector
Span a numeric vector
Gest a numeric vector
Pred a numeric vector
Exp a numeric vector
Danger a numeric vector62
spineMiss
Source
Allison, T. and Chichetti, D. (1976) Sleep in mammals: ecological and constitutional correlates.
Science 194 (4266), 732–734.
The data set was imported from GGobi.

#### Examples 
```{r}
data(sleep, package = "VIM")
summary(sleep)
aggr(sleep)
spineMiss
Spineplot with information about missing/imputed values

#### Description 

Spineplot or spinogram with highlighting of missing/imputed values in other variables by splitting
each cell into two parts. Additionally, information about missing/imputed values in the variable of
interest is shown on the right hand side.

#### Usage 
<pre><code>

spineMiss(x, delimiter = NULL, pos = 1, selection = c("any", "all"),
breaks = "Sturges", right = TRUE, col = c("skyblue", "red", "skyblue4",
"red4", "orange", "orange4"), border = NULL, main = NULL, sub = NULL,
xlab = NULL, ylab = NULL, axes = TRUE, labels = axes,
only.miss = TRUE, miss.labels = axes, interactive = TRUE, ...)
Arguments
x a vector, matrix or data.frame.
delimiter a character-vector to distinguish between variables and imputation-indices for imputed variables (therefore, x needs to have colnames). If given, it is used
to determine the corresponding imputation-index for any imputed variable (a logical-vector indicating which values of the variable have been imputed). If
such imputation-indices are found, they are used for highlighting and the colors are adjusted according to the given colors for imputed variables (see col).
pos a numeric value giving the index of the variable of interest. Additional variables
in x are used for highlighting.
selection the selection method for highlighting missing/imputed values in multiple addi-
tional variables. Possible values are "any" (highlighting of missing/imputed val-
ues in any of the additional variables) and "all" (highlighting of missing/imputed values in all of the additional variables).spineMiss
63
breaks if the variable of interest is numeric, breaks controls the breakpoints (see hist
for possible values).
right logical; if TRUE and the variable of interest is numeric, the spinogram cells are
right-closed (left-open) intervals.
col a vector of length six giving the colors to be used. If only one color is supplied, the bars are transparent and the supplied color is used for highlighting
missing/imputed values. Else if two colors are supplied, they are recycled.
border the color to be used for the border of the cells. Use border=NA to omit borders.
main, sub main and sub title.
xlab, ylab axis labels.
axes a logical indicating whether axes should be drawn on the plot.
labels if the variable of interest is categorical, either a logical indicating whether labels should be plotted below each cell, or a character vector giving the labels. This
is ignored if the variable of interest is numeric.
only.miss logical; if TRUE, the missing/imputed values in the variable of interest are also visualized by a cell in the spineplot or spinogram. Otherwise, a small spineplot
is drawn on the right hand side (see ‘
#### Details 
’).
miss.labels either a logical indicating whether label(s) should be plotted below the cell(s)
on the right hand side, or a character string or vector giving the label(s) ((see details)..
interactive a logical indicating whether the variables can be switched interactively (see ‘De-
tails’).
... further graphical parameters to be passed to title and axis.

#### Details 

A spineplot is created if the variable of interest is categorial and a spinogram if it is numerical.
The horizontal axis is scaled according to relative frequencies of the categories/classes. If more
than one variable is supplied, the cells are split according to missingness/number of imputed values
in the additional variables. Thus the proportion of highlighted observations in each category/class
is displayed on the vertical axis. Since the height of each cell corresponds to the proportion of
highlighted observations, it is now possible to compare the proportions of missing/imputed values
among the different categories/classes.
If only.miss=TRUE, the missing/imputed values in the variable of interest are also visualized by a
cell in the spine plot or spinogram. If additional variables are supplied, this cell is again split into
two parts according to missingness/number if imputed values in the additional variables.
Otherwise, a small spineplot that visualizes missing/imputed values in the variable of interest is
drawn on the right hand side. The first cell corresponds to observed values and the second cell to
missing/imputed values. Each of the two cells is again split into two parts according to missing-
ness/number of imputed values in the additional variables. Note that this display does not make
sense if only one variable is supplied, therefore only.miss is ignored in that case.
If interactive=TRUE, clicking in the left margin of the plot results in switching to the previous
variable and clicking in the right margin results in switching to the next variable. Clicking anywhere
else on the graphics device quits the interactive session.64
tao
Value
a table containing the frequencies corresponding to the cells.
Note
Some of the argument names and positions have changed with version 1.3 due to extended func-
tionality and for more consistency with other plot functions in VIM. For back compatibility, the
arguments xaxlabels and missaxlabels can still be supplied to ...{} and are handled correctly.
Nevertheless, they are deprecated and no longer documented. Use labels and miss.labels in-
stead.
The code is based on the function spineplot by Achim Zeileis.
Author(s)
Andreas Alfons, Matthias Templ, modifications by Bernd Prantner
References
M. Templ, A. Alfons, P. Filzmoser (2012) Exploring incomplete data using visualization tools.
Journal of Advances in Data Analysis and Classification, Online first. DOI: 10.1007/s11634-011-
0102-y.
See Also
histMiss, barMiss, mosaicMiss

#### Examples 
```{r}
data(tao, package = "VIM")
data(sleep, package = "VIM")
## for missing values
spineMiss(tao[, c("Air.Temp", "Humidity")])
spineMiss(sleep[, c("Exp", "Sleep")])
## for imputed values
spineMiss(kNN(tao[, c("Air.Temp", "Humidity")]), delimiter = "_imp")
spineMiss(kNN(sleep[, c("Exp", "Sleep")]), delimiter = "_imp")
tao
Tropical Atmosphere Ocean (TAO) project data

#### Description 

A small subsample of the Tropical Atmosphere Ocean (TAO) project data, derived from the GGOBI
project.testdata
65
Format
A data frame with 736 observations on the following 8 variables.
Year a numeric vector
Latitude a numeric vector
Longitude a numeric vector
Sea.Surface.Temp a numeric vector
Air.Temp a numeric vector
Humidity a numeric vector
UWind a numeric vector
VWind a numeric vector

#### Details 

All cases recorded for five locations and two time periods.
Source
http://www.pmel.noaa.gov/tao/

#### Examples 
```{r}
data(tao, package = "VIM")
summary(tao)
aggr(tao)
testdata
Simulated data set for testing purpose

#### Description 

2 numeric, 2 binary, 2 nominal and 2 mixed (semi-continous) variables
Format
The format is: List of 4 $ wna :’data.frame’: 500 obs. of 8 variables: ..$ x1: num [1:500] 10.87
9.53 7.83 8.53 8.67 ... ..$ x2: num [1:500] 10.9 9.32 7.68 8.2 8.41 ... ..$ c1: Factor w/ 4 levels
"a","b","c","d": 3 2 2 1 2 2 1 3 3 2 ... ..$ c2: Factor w/ 4 levels "a","b","c","d": 2 3 2 2 2 2 2 4 2 2
... ..$ b1: Factor w/ 2 levels "0","1": 2 2 1 2 1 2 1 2 1 1 ... ..$ b2: Factor w/ 2 levels "0","1": 2 2 1 1
1 1 1 2 2 2 ... ..$ m1: num [1:500] 0 8.29 9.08 0 0 ... ..$ m2: num [1:500] 10.66 9.39 7.8 8.11 7.33
... $ wona :’data.frame’: 500 obs. of 8 variables: ..$ x1: num [1:500] 10.87 9.53 7.83 8.53 8.67 ...
..$ x2: num [1:500] 10.9 9.32 7.68 8.2 8.41 ... ..$ c1: Factor w/ 4 levels "a","b","c","d": 3 2 2 1 2 2
1 3 3 2 ... ..$ c2: Factor w/ 4 levels "a","b","c","d": 2 3 2 2 2 2 2 4 2 2 ... ..$ b1: Factor w/ 2 levels
"0","1": 2 2 1 2 1 2 1 2 1 1 ... ..$ b2: Factor w/ 2 levels "0","1": 2 2 1 1 1 1 1 2 2 2 ... ..$ m1: num
[1:500] 0 8.29 9.08 0 0 ... ..$ m2: num [1:500] 10.66 9.39 7.8 8.11 7.33 ... $ mixed : chr [1:2] "m1"
"m2" $ outlierInd: NULL66
vmGUIenvir

#### Examples 
```{r}
data(testdata)
vmGUIenvir
Environment for the GUI for Visualization and Imputation of Missing
Values

#### Description 

Location were everything from package VIM and VIMGUI is stored.

#### Usage 
<pre><code>

vmGUIenvir
putVm(x, value)
getVm(x, mode = "any")
existsVm(x, mode = "any")
rmVm(...)
Arguments
x object name
value value to be assigned to x
mode see ’exists’
... see ’rm’
Format
An object of class environment of length 0.

#### Details 

Internal information regarding the VIM GUI is stored in the environment vmGUIenvir.
Author(s)
Andreas Alfons, based on an initial design by Matthias Templ, modifications by Bernd PrantnervmGUIenvir
67
References
M. Templ, A. Alfons, P. Filzmoser (2012) Exploring incomplete data using visualization tools.
Journal of Advances in Data Analysis and Classification, Online first. DOI: 10.1007/s11634-011-
0102-y.
