Package ‘pivottabler’
February 17, 2019
Type Package
Title Create Pivot Tables in R
Version 1.1.0

#### Description
 Create regular pivot tables with just a few lines of R.
More complex pivot tables can also be created, e.g. pivot tables
with irregular layouts, multiple calculations and/or derived
calculations based on multiple data frames. Pivot tables are
constructed using R only and can be written to a range of
output formats (plain text, 'HTML', 'Latex' and 'Excel'),
including with styling/formatting.
Depends R (>= 3.3.0)
Imports R6 (>= 2.2.0), dplyr (>= 0.5.0), data.table (>= 1.10.0),
jsonlite (>= 1.1), htmltools(>= 0.3.5), htmlwidgets (>= 0.8)
Suggests ggplot2 (>= 2.2.0), lubridate (>= 1.5.0), listviewer (>=
1.4.0), openxlsx (>= 4.0.17), basictabler (>= 0.2.0), shiny,
knitr, rmarkdown, testthat
License GPL-3
Encoding UTF-8
LazyData true
URL https://github.com/cbailiss/pivottabler
BugReports https://github.com/cbailiss/pivottabler/issues
VignetteBuilder knitr
RoxygenNote 6.1.1
NeedsCompilation no
Author Christopher Bailiss [aut, cre]
Maintainer Christopher Bailiss <cbailiss@gmail.com>
Repository CRAN
Date/Publication 2019-02-17 13:50:03 UTC
1R topics documented:
2
R topics documented:
bhmtraindisruption . . . . . . . .
bhmtrains . . . . . . . . . . . . .
checkArgument . . . . . . . . . .
cleanCssValue . . . . . . . . . . .
containsText . . . . . . . . . . . .
convertPvtStyleToBasicStyle . . .
convertPvtTblToBasicTbl . . . . .
exportValueAs . . . . . . . . . . .
getBlankTheme . . . . . . . . . .
getCompactTheme . . . . . . . .
getDefaultTheme . . . . . . . . .
getLargePlainTheme . . . . . . .
getNextPosition . . . . . . . . . .
getPvtStyleDeclarations . . . . . .
getSimpleColoredTheme . . . . .
getTheme . . . . . . . . . . . . .
getXlBorderFromCssBorder . . .
getXlBorderStyleFromCssBorder .
isNumericValue . . . . . . . . . .
isTextValue . . . . . . . . . . . .
parseColor . . . . . . . . . . . . .
parseCssBorder . . . . . . . . . .
parseCssSizeToPt . . . . . . . . .
parseCssSizeToPx . . . . . . . . .
parseCssString . . . . . . . . . .
parseXlBorder . . . . . . . . . . .
PivotBatch . . . . . . . . . . . . .
PivotBatchCalculator . . . . . . .
PivotBatchStatistics . . . . . . . .
PivotCalculation . . . . . . . . . .
PivotCalculationGroup . . . . . .
PivotCalculationGroups . . . . . .
PivotCalculator . . . . . . . . . .
PivotCell . . . . . . . . . . . . .
PivotCells . . . . . . . . . . . . .
PivotData . . . . . . . . . . . . .
PivotDataGroup . . . . . . . . . .
PivotFilter . . . . . . . . . . . . .
PivotFilterOverrides . . . . . . . .
PivotFilters . . . . . . . . . . . .
PivotHtmlRenderer . . . . . . . .
PivotLatexRenderer . . . . . . . .
PivotOpenXlsxRenderer . . . . .
PivotOpenXlsxStyle . . . . . . . .
PivotOpenXlsxStyles . . . . . . .
PivotStyle . . . . . . . . . . . . .
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
3
4
5
6
6
7
7
8
8
9
9
10
10
11
11
12
12
13
13
14
14
15
15
16
16
17
17
18
19
20
22
23
24
25
27
28
29
31
32
34
35
36
37
38
40
41bhmtraindisruption
PivotStyles . . . .
PivotTable . . . . .
pivottabler . . . . .
pivottablerOutput .
processIdentifier . .
processIdentifiers .
pvtperfresults . . .
pvtperfsummary . .
qhpvt . . . . . . .
qlpvt . . . . . . . .
qpvt . . . . . . . .
renderBasicTable .
renderPivottabler .
skipExportingValue
trainstations . . . .
3
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
.
Index
bhmtraindisruption
42
43
47
48
48
49
49
50
51
52
52
54
54
55
55
56
Birmingham Train Disruptions, Dec 2016-Feb 2017.

#### Description

A dataset containing details of the trains in the bhmtrains dataset that were partially/wholly can-
celled and/or reinstated.

#### Usage
<pre><code>

bhmtraindisruption
Format
A data frame with 2982 rows and 10 variables:
ServiceId Unique train identifier
LastCancellationAt Time of the last cancellation
LastCancellationLocation 3-letter code denoting the location of the last cancellation
LastCancellationReasonCategory The broad reason why the train was cancelled
LastCancellationReasonDesc A more specific reason why the train was cancelled
LastReinstatedAt Time of the last reinstatement
LastChangeOfOriginAt Time of the last change of origin
LastChangeOfOriginLocation 3-letter code denoting the location of the last change of origin
LastChangeOfOriginReasonCategory The broad reason why the origin was changed
LastChangeOfOriginReasonDesc A more specific reason why the origin was changed
Source
http://www.recenttraintimes.co.uk/4
bhmtrains
bhmtrains
Birmingham Trains, Dec 2016-Feb 2017.

#### Description

A dataset containing all of the trains that either originated at, passed through or terminated at Birm-
ingham New Street railway station in the UK between 1st December 2016 and 28th February 2017
inclusive.

#### Usage
<pre><code>

bhmtrains
Format
A data frame with 83710 rows and 16 variables:
ServiceId Unique train identifier
Status Train status: A=Active, C=Cancelled, R=Reinstated
TOC Train operating company
TrainCategory Express Passenger or Ordinary Passenger
PowerType DMU=Diesel Multiple Unit, EMU=Electrical Multiple Unit, HST=High Speed Train
SchedSpeedMPH Scheduled maximum speed (in miles per hour)
Origin 3-letter code denoting the scheduled origin of the train
OriginGbttDeparture Scheduled departure time in the Great Britain Train Timetable (GBTT)
from the origin station
OriginActualDeparture Actual departure time from the origin station
GbttArrival Scheduled arrival time in Birmingham in the GBTT
ActualArrival Actual arrival time in Birmingham in the GBTT
GbttDeparture Scheduled departure time from Birmingham in the GBTT
ActualDeparture Actual departure time from Birmingham in the GBTT
Destination 3-letter code denoting the scheduled destination of the train
DestinationGbttArrival Scheduled arrival time in the GBTT at the destination station
DestinationActualArrival Actual arrival time at the destination station
Source
http://www.recenttraintimes.co.uk/checkArgument
checkArgument
5
Perform basic checks on a function argument.

#### Description

checkArgument is a utility function that provides basic assurances about function argument values
and generates standardised error messages when invalid values are encountered.

#### Usage
<pre><code>

checkArgument(argumentCheckMode, checkDataTypes, className, methodName,
argumentValue, isMissing, allowMissing = FALSE, allowNull = FALSE,
allowedClasses = NULL, mustBeAtomic = FALSE,
allowedListElementClasses = NULL, listElementsMustBeAtomic = FALSE,
allowedValues = NULL, minValue = NULL, maxValue = NULL,
maxLength = NULL)

</code></pre>
#### Arguments

argumentCheckMode
A number between 0 and 4 specifying the checks to perform.
checkDataTypes A logical value specifying whether the data types should be checked when ar-
gumentCheckMode=3
className The name of the calling class, for inclusion in error messages.
methodName The name of the calling method, for inclusion in error messages.
argumentValue The value to check.
isMissing Whether the argument is missing in the calling function.
allowMissing Whether missing values are permitted.
allowNull Whether null values are permitted.
allowedClasses The names of the allowed classes for argumentValue.
mustBeAtomic
Whether the argument value must be atomic.
allowedListElementClasses
For argument values that are lists(), the names of the allowed classes for the
elements in the list.
listElementsMustBeAtomic
For argument values that are lists(), whether the list elements must be atomic.
allowedValues For argument values that must be one value from a set list, the list of allowed
values.
minValue For numerical values, the lowest allowed value.
maxValue For numerical values, the highest allowed value.
maxLength For character values, the maximum allowed length (in characters) of the value.6
containsText
Value
No return value. If invalid values are encountered, the stop() function is used to interrupt execu-
tion.
cleanCssValue
Cleans up a CSS attribute value.

#### Description

cleanCssValue is a utility function that performs some basic cleanup on CSS attribute values.
Leading and trailing whitespace is removed. The CSS values "initial" and "inherit" are blocked.
The function is vectorised so can be used with arrays.

#### Usage
<pre><code>

cleanCssValue(cssValue)

</code></pre>
#### Arguments

cssValue
The value to cleanup.
Value
The cleaned value.
containsText
Check whether a text value is present in another text value.

#### Description

containsText is a utility function returns TRUE if one text value is present in another. Case
sensitive. If textToSearch is a vector, returns TRUE if any element contains textToFind.

#### Usage
<pre><code>

containsText(textToSearch, textToFind)

</code></pre>
#### Arguments

textToSearch The value to be searched.
textToFind The value to find.
Value
TRUE if the textToFind value is found.convertPvtStyleToBasicStyle
7
convertPvtStyleToBasicStyle
Convert a pivot table style to a basictabler style.

#### Description

convertPvtStyleToBasicStyle is a utility function that converts a pivot table style to a basictabler
style from the basictabler package.

#### Usage
<pre><code>

convertPvtStyleToBasicStyle(btbl = NULL, pvtStyle = NULL)

</code></pre>
#### Arguments

btbl The basic table that will own the new style.
pvtStyle The pivot style to convert.
Value
a basictabler style.
convertPvtTblToBasicTbl
Convert a pivot table to a basic table.

#### Description

convertPvtTblToBasicTbl is a utility function that converts a pivot table to a basic table from the
basictabler package.

#### Usage
<pre><code>

convertPvtTblToBasicTbl(pvt = NULL, exportOptions = NULL)

</code></pre>
#### Arguments

pvt The pivot table to convert.
exportOptions Options specifying how values are exported.
Value
a basictabler table.8
getBlankTheme
Replace the current value with a placeholder during export.
exportValueAs

#### Description

exportValueAs is a utility function that returns either the original value or a replacement place-
holder value for export.

#### Usage
<pre><code>

exportValueAs(rawValue, formattedValue, exportOptions,
blankValue = character(0))

</code></pre>
#### Arguments

rawValue
The raw value to check.
formattedValue The formatted value to be exported.
exportOptions A list of options controlling export behaviour.
blankValue The ’placeholder’ value to be exported when skipping the value.
Value
Either the original value or a placeholder value.
Get an empty theme for applying no styling to a table.
getBlankTheme

#### Description

Get an empty theme for applying no styling to a table.

#### Usage
<pre><code>

getBlankTheme(parentPivot, themeName = "blank")

</code></pre>
#### Arguments

parentPivot Owning pivot table.
themeName The name to use as the new theme name.
Value
A TableStyles object.getCompactTheme
9
Get the compact theme for styling a pivot table.
getCompactTheme

#### Description

Get the compact theme for styling a pivot table.

#### Usage
<pre><code>

getCompactTheme(parentPivot, themeName = "compact")

</code></pre>
#### Arguments

parentPivot Owning pivot table.
themeName The name to use as the new theme name.
Value
A PivotStyles object.
Get the default theme for styling a pivot table.
getDefaultTheme

#### Description

Get the default theme for styling a pivot table.

#### Usage
<pre><code>

getDefaultTheme(parentPivot, themeName = "default")

</code></pre>
#### Arguments

parentPivot Owning pivot table.
themeName The name to use as the new theme name.
Value
A PivotStyles object.10
getNextPosition
getLargePlainTheme
Get the large plain theme for styling a pivot table.

#### Description

Get the large plain theme for styling a pivot table.

#### Usage
<pre><code>

getLargePlainTheme(parentPivot, themeName = "largeplain")

</code></pre>
#### Arguments

parentPivot Owning pivot table.
themeName The name to use as the new theme name.
Value
A PivotStyles object.
getNextPosition
Find the first value in an array that is larger than the specified value.

#### Description

getNextPosition is a utility function that helps when parsing strings that contain delimiters.

#### Usage
<pre><code>

getNextPosition(positions, afterPosition)

</code></pre>
#### Arguments

positions An ordered numeric vector.
afterPosition The value to start searching after.
Value
The first value in the array larger than afterPosition.getPvtStyleDeclarations
11
getPvtStyleDeclarations
Get pivot table style declarations from a pivot table style.

#### Description

getPvtStyleDeclarations is a utility function that reads the styles from a pivot table style.

#### Usage
<pre><code>

getPvtStyleDeclarations(pvtStyle = NULL)

</code></pre>
#### Arguments

pvtStyle
The pivot table style to read.
Value
a list of style declarations.
getSimpleColoredTheme Get a simple coloured theme.

#### Description

Get a simple coloured theme that can be used to style a pivot table into a custom colour scheme.

#### Usage
<pre><code>

getSimpleColoredTheme(parentPivot, themeName = "coloredTheme", colors,
fontName)

</code></pre>
#### Arguments

parentPivot Owning pivot table.
themeName The name to use as the new theme name.
colors The set of colours to use when generating the theme (see the Styling vignette for
details).
fontName The name of the font to use, or a comma separated list (for font-fall-back).
Value
A PivotStyles object.12
getXlBorderFromCssBorder
Get a built-in theme for styling a pivot table.
getTheme

#### Description

getTheme returns the specified theme.

#### Usage
<pre><code>

getTheme(parentPivot, themeName = NULL)

</code></pre>
#### Arguments

parentPivot Owning pivot table.
themeName The name of the theme to retrieve.
Value
A PivotStyles object.
getXlBorderFromCssBorder
Convert CSS border values to those used by the openxlsx package.

#### Description

getXlBorderFromCssBorder parses the CSS combined border declarations (i.e. border, border-
left, border-right, border-top, border-bottom) and returns a list containing an openxlsx border style
and color as separate elements.

#### Usage
<pre><code>

getXlBorderFromCssBorder(text)

</code></pre>
#### Arguments

text
The border declaration to parse.
Value
A list containing two elements: style and color.getXlBorderStyleFromCssBorder
13
getXlBorderStyleFromCssBorder
Convert CSS border values to those used by the openxlsx package.

#### Description

getXlBorderStyleFromCssBorder takes border parameters expressed as a list (containing ele-
ments: width and style) and returns a border style that is compatible with the openxlsx package.

#### Usage
<pre><code>

getXlBorderStyleFromCssBorder(border)

</code></pre>
#### Arguments

border
A list containing elements width and style.
Value
An openxlsx border style.
Check whether a numeric value is present.
isNumericValue

#### Description

isNumericValue is a utility function returns TRUE only when a numeric value is present. NULL,
NA, numeric(0) and integer(0) all return FALSE.

#### Usage
<pre><code>

isNumericValue(value)

</code></pre>
#### Arguments

value
The value to check.
Value
TRUE if a numeric value is present.14
parseColor
Check whether a text value is present.
isTextValue

#### Description

isTextValue is a utility function returns TRUE only when a text value is present. NULL, NA,
character(0) and "" all return FALSE.

#### Usage
<pre><code>

isTextValue(value)

</code></pre>
#### Arguments

value
The value to check.
Value
TRUE if a non-blank text value is present.
Convert a CSS colour into a hex based colour code.
parseColor

#### Description

parseColor converts a colour value specified in CSS to a hex based colour code. Example sup-
ported input values/formats/named colours are: #0080FF, rgb(0, 128, 255), rgba(0, 128, 255, 0.5)
and red, green, etc.

#### Usage
<pre><code>

parseColor(color)

</code></pre>
#### Arguments

color
The colour to convert.
Value
The colour as a hex code, e.g. #FF00A0.parseCssBorder
15
Parse a CSS border value.
parseCssBorder

#### Description

parseCssBorder parses the CSS combined border declarations (i.e. border, border-left, border-
right, border-top, border-bottom) and returns a list containing the width, style and color as separate
elements.

#### Usage
<pre><code>

parseCssBorder(text)

</code></pre>
#### Arguments

text
The border declaration to parse.
Value
A list containing three elements: width, style and color.
parseCssSizeToPt
Convert a CSS size value into points.

#### Description

parseCssSizeToPt will take a CSS style and convert it to points. Supported input size units are in,
cm, mm, pt, pc, px, em, are converted exactly: in, cm, mm, pt, pc: using 1in = 2.54cm = 25.4mm =
72pt = 6pc. The following are converted approximately: px, em, approx 1em=16px=12pt and 100

#### Usage
<pre><code>

parseCssSizeToPt(size)

</code></pre>
#### Arguments

size
A size specified in common CSS units.
Value
The size converted to points.16
parseCssString
parseCssSizeToPx
Convert a CSS size value into pixels

#### Description

parseCssSizeToPx will take a CSS style and convert it to pixels Supported input size units are in,
cm, mm, pt, pc, px, em, are converted exactly: in, cm, mm, pt, pc: using 1in = 2.54cm = 25.4mm =
72pt = 6pc. The following are converted approximately: px, em, approx 1em=16px=12pt and 100

#### Usage
<pre><code>

parseCssSizeToPx(size)

</code></pre>
#### Arguments

size
A size specified in common CSS units.
Value
The size converted to pixels.
parseCssString
Split a CSS attribute value into a vector/array.

#### Description

parseCssString is a utility function that splits a string into a vector/array. The function pays
attention to text qualifiers (single and double quotes) so won’t split if the delimiter occurs inside a
value.

#### Usage
<pre><code>

parseCssString(text, separator = ",", removeEmptyString = TRUE)

</code></pre>
#### Arguments

text
The text to split.
separator
The field separator, default comma.
removeEmptyString
TRUE to not return empty string / whitespace values.
Value
An R vector containing the values from text split up.parseXlBorder
17
Parse an xl-border value.
parseXlBorder

#### Description

parseXlBorder parses the combined xl border declarations (i.e. xl-border, xl-border-left, xl-
border-right, xl-border-top, xl-border-bottom) and returns a list containing style and color as sepa-
rate elements.

#### Usage
<pre><code>

parseXlBorder(text)

</code></pre>
#### Arguments

text
The border declaration to parse.
Value
A list containing two elements: style and color.
PivotBatch
A class that represents a batch calculation.

#### Description

The PivotBatch class represents one combination of variables and calculations that are needed when
calculating the values of cells in a pivot table.

#### Usage
<pre><code>

PivotBatch
Format
R6Class object.
Value
Object of R6Class with properties and methods that help to (do xyz).18
PivotBatchCalculator
Fields
parentPivot Owning pivot table.
batchId A unique integer identifier for this batch.
compatibleCount The number of pivot cell calculations that this batch supports.
evaluated TRUE if this batch has been evaluated.
results The results (a data frame) of the evaluation of this batch.
asString A text description of this batch.
Methods
Documentation For more complete explanations and examples please see the extensive vignettes
supplied with this package.
new(...) Create a new Pivot Batch.
isCompatible(dataName=NULL, variableNames=NULL) Checks whether the specified data name
and combination of variable names are compatible with this batch.
addCompatible(values=NULL, calculationName=NULL, calculationGroupName=NULL) Adds
another cell calculation to this batch.
getCalculationInternalName(calculationName=NULL, calculationGroupName=NULL) Find
the internal name for this calculation.
evaluateBatch() Evaluate this batch.
getSummaryValueFromBatch(filters=NULL, calculationName=NULL, calculationGroupName=NULL)
Retrieve a value from a batch that has already been evaluated.

#### Examples
```{r}

# This class should only be created by the pivot table.
# It is not intended to be created outside of the pivot table.
PivotBatchCalculator
A class that calculates the values for multiple cells.

#### Description

The PivotBatchCalculator class calculates the values for multiple cells in the pivot table in one
evaluation step.

#### Usage
<pre><code>

PivotBatchCalculator
Format
R6Class object.PivotBatchStatistics
19
Value
Object of R6Class with properties and methods that help to perform calculations in batch.
Fields
parentPivot Owning pivot table.
batchCount The number of batches generated for the pivot table.
calculationSummary A summary of the batch compatibility for each calculation.
batchSummary A summary of the batches in the pivot table.
Methods
Documentation For more complete explanations and examples please see the extensive vignettes
supplied with this package.
new(...) Create a new batch calculator.
reset() Clears any batches that currently exist in the batch calculator.
isFiltersBatchCompatible(filters=NULL) Determines whether a set of filters are compatible
with batch calculations.
generateBatchesForNamedCalculationEvaluation1(dataName=NULL, calculationName=NULL, calculationGroupN
Generates one or more batches for a named calculation and single working filters object.
generateBatchesForNamedCalculationEvaluation2(calculationName=NULL, calculationGroupName=NULL, worki
Generates one or more batches for a named calculation and set of working working data asso-
ciated with a cell.
generateBatchesForCellEvaluation() Generates one or batches for a pivot table cell.
evaluateBatches() Evaluates batch calculations.
getSummaryValueFromBatch(batchName=NULL, calculationName=NULL, calculationGroupName=NULL, workingFil
Retrieve a value from a the results of a batch that has been evaluated.

#### Examples
```{r}

# This class should only be created by the pivot table.
# It is not intended to be created outside of the pivot table.
```
