Package ‘lemon’
October 30, 2018
Type Package
Title Freshing Up your 'ggplot2' Plots
URL https://github.com/stefanedwards/lemon
BugReports https://github.com/stefanedwards/lemon/issues
Version 0.4.2
Date 2018-10-30
Description Functions for working with legends and axis lines of 'ggplot2',
facets that repeat axis lines on all panels, and some 'knitr' extensions.
Depends R (>= 3.1.0)
Imports ggplot2 (>= 2.2.0), plyr, grid, gridExtra, gtable, knitr (>=
1.12), lattice, scales
License GPL-3
Encoding UTF-8
LazyData true
RoxygenNote 6.1.0.9000
Collate 'ggplot2.r' 'lemon-plot.r' 'axis-annotation.r' 'brackets.R'
'coord-flex.r' 'coord-capped.r' 'dot.r' 'facet-rep-lab.r'
'facet-wrap.r' 'geom-pointline.r' 'lemon_print.r'
'geom-siderange.r' 'grob_utils.r' 'gtable_show-.r'
'guides-axis.r' 'legends.r' 'lemon.r' 'scale-symmetric.r'
Suggests rmarkdown, stringr, dplyr, testthat
VignetteBuilder knitr
NeedsCompilation no
Author Stefan McKinnon Edwards [aut, ctb, cre]
(<https://orcid.org/0000-0002-4628-8148>),
Baptiste Auguie [ctb] (For g_legend and grid_arrange_shared_legend),
Shaun Jackman [ctb] (For grid_arrange_shared_legend),
Hadley Wickham [ctb] (ggplot2 functions),
Winston Chang [ctb] (ggplot2 functions)
Maintainer Stefan McKinnon Edwards <sme@iysik.com>
12
.dot
Repository CRAN
Date/Publication 2018-10-30 20:50:03 UTC
R topics documented:
.dot . . . . . . . . . . . . .
annotate_y_axis . . . . . . .
brackets_horizontal . . . . .
coord_capped_cart . . . . .
coord_flex_cart . . . . . . .
element_render . . . . . . .
facet_rep_grid . . . . . . . .
geom_pointpath . . . . . . .
geom_siderange . . . . . . .
get_panel_range . . . . . . .
grid_arrange_shared_legend
gtable_show_grill . . . . . .
guidebox_as_column . . . .
g_legend . . . . . . . . . . .
is.small . . . . . . . . . . .
lemon . . . . . . . . . . . .
lemon_print . . . . . . . . .
remove_labels_from_axis . .
reposition_legend . . . . . .
scale_x_symmetric . . . . 




























































































.
.
.
Index
.dot













.
.
2
3
5
6
7
9
10
10
13
15
16
17
18
19
20
21
22
23
24
25
27







g_legend
Extract ggplot legends
Description
Extracts the legend (’guide-box’) from a ggplot2 object.
Usage
g_legend(a.gplot)
Arguments
a.gplot
ggplot2 or gtable object.
Details
The extraction is applied after the plot is trained and themes are applied. Modifying the legend is
easiest by applying themes etc. to the ggplot2 object, before calling g_legend.
An alternative method for extracting the legend is using gtable::gtable_filter:
gtable_filter(ggplotGrob(a.ggplot.obj), 'guide-box')
This method however returns a gtable object which encapsulates the entire legend. The legend
itself may be a collection of gtable. We have only noticed a problem with this extra layer when
using the returned legend with arrangeGrob (see examples).
Value
gtable (grob) object. Draw with grid.draw.
Author(s)
Baptiste Auguié
See Also
grid_arrange_shared_legend, reposition_legend, gtable_filter20
is.small
Examples
library(ggplot2)
library(gtable)
library(grid)
library(gridExtra)
library(gtable)
dsamp <- diamonds[sample(nrow(diamonds), 1000), ]
(d <- ggplot(dsamp, aes(carat, price)) +
geom_point(aes(colour = clarity)) +
theme(legend.position='bottom'))
legend <- g_legend(d)
grid.newpage()
grid.draw(legend)
(d2 <- ggplot(dsamp, aes(x=carat, fill=clarity)) +
geom_histogram(binwidth=0.1) +
theme(legend.position='bottom'))
grid.arrange(d + theme(legend.position='hidden'),
d2 + theme(legend.position='hidden'),
bottom=legend$grobs[[1]])
# Above fails with more than one guide
legend2 <- gtable_filter(ggplotGrob(d), 'guide-box')
grid.arrange(d + theme(legend.position='hidden'),
d2 + theme(legend.position='hidden'),
bottom=legend2$grobs[[1]]$grobs[[1]])
# Above fails with more than one guide
is.small
Is a given unit ’small’?
Description
Uses a holistic approach to determine whether a unit is ’small’, i.e. less than 1 cm, 1 line, 10 pt, or
0.4 in.
Usage
is.small(x)
Arguments
x
A unit.lemon
21
Details
Based on arbitraily chosen definitions of ’small’, this function can return TRUE or FALSE if a unit is
’small’.
So far, less than 1 cm, 1 line, 10 pt, or 0.4 inches is defined as being ’small’. Unresolved sizes,
suchs as ’grobheight’, ’grobwidth’, or ’null’ are not small. Units based on arithmetic, such as sum
of multiple units, are also not small. NAs are returned for undecided sizes.
Value
Logical or NA.
lemon
Freshing up your ggplots
Description
Collection of misc. functions for changing subtle aspects of ggplots. Works mostly on gtables
produced prior to printing.
Functions for axis
See coord_capped_cart and coord_flex_cart. The latter is a shorthand version of the former. It
automatically uses capped_horizontal and capped_vertical, but both accepts these as well as
brackets_horizontal and brackets_vertical.
Legends
Extract legend g_legend
Many plots, one legend grid_arrange_shared_legend
Place legend exactly on plot reposition_legend
Facets
facet_rep_grid and facet_rep_wrap are extensions to the wellknown facet_grid and facet_wrap
where axis lines and labels are drawn on all panels.
Extending knitr
We automatically load knitr’s knit_print for data frames and dplyr tables to provide automatic
pretty printing of data frame using kable.
See lemon_print or vignette('lemon_print', 'lemon').
Relative paths safe from hanging directory: .dot.22
lemon_print
Author(s)
Stefan McKinnon Edwards <sme@iysik.com>
Contributions from Baptiste Auguié on g_legend and grid_arrange_shared_legend.
Contributions from Shaun Jackman on grid_arrange_shared_legend.
Source
https://github.com/stefanedwards/lemon
See Also
Useful links:
• https://github.com/stefanedwards/lemon
• Report bugs at https://github.com/stefanedwards/lemon/issues
lemon_print
knitr extension: Always use ‘kable‘ for data frames.
Description
Convenience function for working with R Notebooks that ensures data frames (and dplyr tables) are
printed with kable while allowing RStudio to render the data frame dynamically for inline display.
Usage
lemon_print(x, options, ...)
## S3 method for class 'data.frame'
lemon_print(x, options, ...)
## S3 method for class 'table'
lemon_print(x, options, ...)
Arguments
x an data frame or dplyr table object to be printed
options Current chunk options are passed through this argument.
... Ignored for now.
Details
These functions divert data frame and summary output to kable for nicely printing the output.
For options to kable, they can be given directly as chunk-options (see arguments to kable), or
though as a list to a special chunk-option kable.opts.
For more examples, see vignette('lemon_print', package='lemon').remove_labels_from_axis
23
Knitr usage
To use for a single chunk, do
```{r render=lemon_print,caption='My data frame'}
data.frame
```
Note: We are not calling the function, but instead refering to it.
An alternate route for specifying kable arguments is as:
```{r render=lemon_print,kable.opts=list(align='l')}
data.frame
```
The option kable.opts takes precendence over arguments given directly as chunk-options.
To enable as default printing method for all chunks, include
knit_print.data.frame <- lemon_print
knit_print.table <- lemon_print
knit_print.grouped_df <- lemon_print # enableds dplyr results
knit_print.tibble <- lemon_print
knit_print.tbl <- lemon_print
Note: We are not calling the function, but instead assigning the knit_print functions for some
classes.
To disable, temporarily, specify chunk option:
```{r render=normal_print}`
data.frame
```
See Also
knit_print, kable
remove_labels_from_axis
Removes labels from axis grobs.
Description
Called from FacetGridRepeatLabels.
Usage
remove_labels_from_axis(axisgrob)24
reposition_legend
Arguments
axisgrob
reposition_legend
Grob with an axis.
Reposition a legend onto a panel
Description
Repositions a legend onto a panel, by either taking it from the same ggplot, or by using another.
Works on both ggplot2 and gtable objects, and can accept any grob as legend.
Usage
reposition_legend(aplot, position = NULL, legend = NULL,
panel = "panel", x = NULL, y = NULL, just = NULL,
name = "guide-box", clip = "on", offset = c(0, 0), z = Inf,
plot = TRUE)
Arguments
aplot a ggplot2 or gtable object.
position Where to place the legend in the panel. Overrules just argument.
legend The legend to place, if NULL (default), it is extracted from aplot if this is a
ggplot2 object.
panel Name of panel in gtable. See description.
x horisontal coordinate of legend, with 0 at left.
y vertical coordiante of legend, with 0 at bottom.
just ’Anchor point’ of legend; it is this point of the legend that is placed at the x and
y coordinates.
name, clip, z Parameters forwarded to gtable_add_grob.
offset Numeric vector, sets distance from edge of panel. First element for horisontal
distance, second for vertical. Not used by arguments x and y.
plot Logical, when TRUE (default), draws plot with legend repositioned on a new
page.
Details
To modify the look of the legend, use themes and the natural ggplot functions found in guide_legend.
Positioning is done by argument position which places the panel relative in panel (see below).
position resolves to three variables, x, y, and just. x and y is the coordinate in panel, where the
anchorpoint of the legend (set via just) is placed. In other words, just='bottom right' places
the bottom right corner of the legend at coordinates (x,y).scale_x_symmetric
25
The positioning can be set by argument position alone, which can be further nudged by setting
position, x, and y. Alternatively, manually positioning can be obtained by setting arguments. x,
y, and just.
Panel name is by default panel, but when using facets it typically takes the form panel-{col}-{row},
but not for wrapped facets. Either print result from ggplotGrob or use gtable_show_names to dis-
play all the names of the gtable object.
panel takes multiple names, and will then use these components’ extremes for placing the legend.
If panel is an integer vector of length 2 or 4, these elements are used directly for top-left and
bottom-right coordinates.
Value
gtable object, invisibly, with legend repositioned. Can be drawn with grid.draw.
Author(s)
Stefan McKinnon Edwards <sme@iysik.com>
See Also
g_legend, grid_arrange_shared_legend
Examples
library(ggplot2)
dsamp <- diamonds[sample(nrow(diamonds), 1000), ]
(d <- ggplot(dsamp, aes(carat, price)) +
geom_point(aes(colour = clarity)))
reposition_legend(d + theme(legend.position='bottom'), 'bottom right')
# To change the orientation of the legend, use theme's descriptors.
reposition_legend(d + theme(legend.position='bottom'), 'top left')
# Use odd specifications, here offset the legend with half its height from the bottom.
reposition_legend(d + theme(legend.position='bottom'), x=0.3, y=0, just=c(0, -0.5))
# For using with facets:
reposition_legend(d + facet_grid(.~cut), 'top left', panel = 'panel-3-1')
scale_x_symmetric
Symmetrix position scale for continuous x and y
Description
scale_x_symmetric and scale_y_symmetric are like the default scales for continuous x and y,
but ensures that the resulting scale is centered around mid. Does not work when setting limits on
the scale.26
scale_x_symmetric
Usage
scale_x_symmetric(mid = 0, ...)
scale_y_symmetric(mid = 0, ...)
Arguments
mid Value to center the scale around.
... Values passed on to scale_continuous.
Examples
library(ggplot2)
df <- expand.grid(a=c(-1,0,1), b=c(-1,0,1))
rnorm2 <- function(x,y,n,sdx,sdy) {
if (missing(sdy))
sdy <- sdx
data.frame(a=x,b=y,x=rnorm(n,x,sdx), y=rnorm(n,y,sdy))
}
df <- mapply(rnorm2,df$a, df$b, MoreArgs=list(n=30,sdx=1),SIMPLIFY=FALSE)
df <- do.call(rbind, df)
(p <- ggplot(df, aes(x=x,y=y)) + geom_point() +
facet_grid(a~b, scales='free_x')
)
p + scale_x_symmetric(mid=0)Index
g_legend, 16, 18, 19, 21, 22, 25
geom_path, 10
geom_point, 10
geom_pointline (geom_pointpath), 10
geom_pointpath, 10
geom_pointrangeline (geom_pointpath), 10
geom_rug, 15
geom_siderange, 13
get_panel_params (get_panel_range), 15
get_panel_range, 15
get_panel_x_range (get_panel_range), 15
get_panel_y_range (get_panel_range), 15
ggplotGrob, 17, 18, 25
gpar, 9, 17
grid.draw, 16, 19, 25
grid.grob, 9
grid_arrange_shared_legend, 16, 19, 21,
22, 25
gtable, 18
gtable_add_grob, 24
gtable_filter, 19
gtable_show_grill, 17
gtable_show_layout, 17
gtable_show_names, 25
gtable_show_names (gtable_show_grill),
17
guide_legend, 24
guidebox_as_column, 18
∗Topic interal
remove_labels_from_axis, 23
.dot, 2, 21
.dot2 (.dot), 2
absoluteGrob, 8
aes, 11, 13
aes_, 11, 13
annotate_x_axis (annotate_y_axis), 3
annotate_y_axis, 3
arrangeGrob, 16, 19
arrow, 12, 14
brackets_horisontal
(brackets_horizontal), 5
brackets_horizontal, 5, 6–8, 21
brackets_vertical, 21
brackets_vertical
(brackets_horizontal), 5
capped_horisontal (coord_capped_cart), 6
capped_horizontal, 6–8, 21
capped_horizontal (coord_capped_cart), 6
capped_vertical, 21
capped_vertical (coord_capped_cart), 6
coord_capped_cart, 5, 6, 21
coord_capped_flip (coord_capped_cart), 6
coord_cartesian, 6
coord_flex_cart, 5, 7, 7, 21
coord_flex_fixed (coord_flex_cart), 7
coord_flex_flip, 7
coord_flex_flip (coord_flex_cart), 7
coord_flip, 6
is.small, 20
kable, 21–23
knit_print, 21, 23
element_render, 9
layer, 12, 14
lemon, 21
lemon-package (lemon), 21
lemon_print, 21, 22
facet_grid, 10, 21
facet_rep_grid, 10, 21
facet_rep_wrap, 21
facet_rep_wrap (facet_rep_grid), 10
facet_wrap, 10, 21
plotmath, 4
2728
points, 14
position_jitter, 11, 14
remove_labels_from_axis, 23
render_gpar (element_render), 9
reposition_legend, 16, 19, 21, 24
scale_continuous, 26
scale_x_symmetric, 25
scale_y_symmetric (scale_x_symmetric),
25
theme, 5, 9, 10
unit, 5, 11
INDEX
