
co_table Compute Conditional Tables

#### Desription

For a contingency table in array form, compute a list of conditional tables given some margins.

#### Usage
<pre><code>

co_table(x, margin, collapse = ".")

</code></pre>
#### Arguments

* ``x``: a contingency table in array form.
* `` margin`` :  margin index(es) or corresponding name(s) of the conditioning variables.
collapse character used when collapsing level names (if more than 1 margin is specified).
Details
This is essentially an interface to [ which is more convenient for arrays of arbitrary dimension.
Value
A list of the resulting conditional tables.
Author(s)
Achim Zeileis <Achim.Zeileis@R-project.org>

#### Examples
```{r}

data("HairEyeColor")
co_table(HairEyeColor, 1)
co_table(HairEyeColor, c("Hair", "Eye"))
co_table(HairEyeColor, 1:2, collapse = "")
DanishWelfare 31
DanishWelfare Danish Welfare Study Data

#### Desription

Data from the Danish Welfare Study.

#### Usage
<pre><code>

data("DanishWelfare")

</code></pre>
#### Format 

A data frame with 180 observations and 5 variables.
Freq frequency.
Alcohol factor indicating daily alcohol consumption: less than 1 unit (<1), 1-2 units (1-2) or more
than 2 units (>2). 1 unit is approximately 1 bottle of beer or 4cl 40% alcohol.
Income factor indicating income group in 1000 DKK (0-50, 50-100, 100-150, >150).
Status factor indicating marriage status (Widow, Married, Unmarried).
Urban factor indicating urbanization: Copenhagen (Copenhagen), Suburbian Copenhagen (Sub-
Copenhagen), three largest cities (LargeCity), other cities (City), countryside (Country).

#### Source
E. B. Andersen (1991), The Statistical Analysis of Categorical Data, page 205.
References
E. B. Andersen (1991), The Statistical Analysis of Categorical Data. 2nd edition. Springer-Verlag,
Berlin.

#### Examples
```{r}

data("DanishWelfare")
ftable(xtabs(Freq ~ ., data = DanishWelfare))
32 distplot
distplot Diagnostic Distribution Plots

#### Desription

Diagnostic distribution plots: poissonness, binomialness and negative binomialness plots.

#### Usage
<pre><code>

distplot(x, type = c("poisson", "binomial", "nbinomial"),
size = NULL, lambda = NULL, legend = TRUE, xlim = NULL, ylim = NULL,
conf_int = TRUE, conf_level = 0.95, main = NULL,
xlab = "Number of occurrences", ylab = "Distribution metameter",
gp = gpar(cex = 0.8), lwd=2, gp_conf_int = gpar(lty = 2),
name = "distplot", newpage = TRUE,
pop =TRUE, return_grob = FALSE, ...)

</code></pre>
#### Arguments

x either a vector of counts, a 1-way table of frequencies of counts or a data frame
or matrix with frequencies in the first column and the corresponding counts in
the second column.
type a character string indicating the distribution.
size the size argument for the binomial and negative binomial distribution. If set to
NULL and type is "binomial", then size is taken to be the maximum count. If
set to NULL and type is "nbinomial", then size is estimated from the data.
lambda parameter of the poisson distribution. If type is "poisson" and lambda is specified
a leveled poissonness plot is produced.
legend logical. Should a legend be plotted?
xlim limits for the x axis.
ylim limits for the y axis.
conf_int logical. Should confidence intervals be plotted?
conf_* `` level`` : confidence level for confidence intervals.
main a title for the plot.
xlab a label for the x axis.
ylab a label for the y axis.
gp a "gpar" object controlling the grid graphical parameters of the points.
gp_conf_int a "gpar" object controlling the grid graphical parameters of the confidence intervals.
lwd line width for the fitted line
name name of the plotting viewport.
newpage logical. Should grid.newpage be called before plotting?
distplot 33
pop logical. Should the viewport created be popped?
return_grob logical. Should a snapshot of the display be returned as a grid grob?
... further 
</code></pre>
#### Arguments
 passed to grid.points.
Details
distplot plots the number of occurrences (counts) against the distribution metameter of the specified
distribution. If the distribution fits the data, the plot should show a straight line. See Friendly
(2000) for details.
In these plots, the open points show the observed count metameters; the filled points show the
confidence interval centers, and the dashed lines show the conf_level confidence intervals for
each point.
Value
Returns invisibly a data frame containing the counts (Counts), frequencies (Freq) and other details
of the computations used to construct the plot.
Author(s)
Achim Zeileis <Achim.Zeileis@R-project.org>
References
D. C. Hoaglin (1980), A poissonness plot, The American Statistican, 34, 146–149.
D. C. Hoaglin & J.W. Tukey (1985), Checking the shape of discrete distributions. In D. C. Hoaglin,
F. Mosteller, J. W. Tukey (eds.), Exploring Data Tables, Trends and Shapes, chapter 9. John Wiley
& Sons, New York.
M. Friendly (2000), Visualizing Categorical Data. SAS Institute, Cary, NC.

#### Examples
```{r}

## Simulated data 
#### Examples
```{r}
:
dummy <- rnbinom(1000, size = 1.5, prob = 0.8)
distplot(dummy, type = "nbinomial")
## Real data 
#### Examples
```{r}
:
data("HorseKicks")
data("Federalist")
data("Saxony")
distplot(HorseKicks, type = "poisson")
distplot(HorseKicks, type = "poisson", lambda = 0.61)
distplot(Federalist, type = "poisson")
distplot(Federalist, type = "nbinomial", size = 1)
distplot(Federalist, type = "nbinomial")
distplot(Saxony, type = "binomial", size = 12)
34 doubledecker
